---
title: "Biocreep Simulation"
author: "Amy Ly"
date: "4/12/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# This package was sourced from https://osf.io/37vch/
source ("informedTtest_functions.R") 

library(tidyverse)
library(gridExtra)
library(metR)
```


#Define Custom Functions


```{r}
#corrected CohenH's value
get_CohenH <- function(k1, k2, n1, n2, NImargin){
  
 h <- 2*asin(sqrt(((k1+k2)/(n1+n2)) - (NImargin/2)))-2*asin(sqrt(((k1+k2)/(n1+n2)) + (NImargin/2)))
                                                    
  return(h)
}

#corrected NI BF test

NI_BFprop <- function (n1, n2, k1, k2, NImargin, prior.location, prior.scale)
{
  #modified SE equation to not be pooled proportion
  SE <- sqrt (((k1/n1)*(1-(k1/n1))/n1) + ((k2/n2)*(1-(k2/n2))/n2))
  
  #signs are opposite from the code (higher propor is good for our case)
  Z <- (((k2/n2) - (k1/n1)) + NImargin) / SE
  
  Res <- bf10_t (Z, n1 = n1, n2 = n2, independentSamples = T, prior.location = prior.location, 
                prior.scale =prior.scale, prior.df = 1)
  
  BFmin_plus <- Res[[2]]/Res[[3]]
  
  return(BFmin_plus)
}


#extract a common legend for when we put all instructional plots on a grid
get_only_legend <- function(plot) {
    
  # get tabular interpretation of plot
  plot_table <- ggplot_gtable(ggplot_build(plot)) 
    
  #  Mark only legend in plot
  legend_plot <- which(sapply(plot_table$grobs, function(x) x$name) == "guide-box") 
                              
  # extract legend
  legend <- plot_table$grobs[[legend_plot]]
                              
  # return legend
  return(legend) 
}

```


## Case Study Traits

* Have dichotomous data
* 1 sided test at a significance level of 0.05
* 2-sided 90% CI
* power = 0.8 and is based on estimation that mortality rates are 5% in study population
* sample size calculations require 650 patients in each arm
* 90-day all-cause mortality determined by random-effects logistic regression model
* Test for NI: the upper limit of 95% CI of estimated mortality in each group should not be more than 3% above the point estimate of the group with the lowest mortality
* Test for difference: LR test of model with and without antibiotic treatment
* NI figures based on risk differences of the treatment strategy compared to control arm


Given the following information:

```{r}

n1 <- 656  # n for group 1 (beta-lactam)
n2 <- 739  # n for group 2 (beta-lactam-macrolide)
n3 <- 888 # n for group 1 (fluoroquinolone)
k1 <- 59  # mortality count in group 1 (beta-lactam)
k2 <- 82  # mortality count in group 2 (beta-lactam-macrolide)
k3 <- 78  # mortality count in group 1 (fluoroquinolone)

NImargin <- 0.03   # Non-inferiority margin in proportion points

```


Bayes Factor Analysis for beta-lactam vs beta-lactam-fluoroquinone


```{r}

SE <- sqrt (((k1/n1)*(1-(k1/n1))/n1) + ((k3/n3)*(1-(k3/n3))/n3))
  
Z <- (((k3/n3) - (k1/n1)) + NImargin) / SE

xlim <- range(c(-1, 1))
xx <- seq(xlim[1], xlim[2], length.out = 400)
yy <- posterior_t(delta = xx, t = Z, n1 = n3, n2 = n1,
                  independentSamples = TRUE,
                  prior.location = -get_CohenH(k1, k3, n1, n3, NImargin), 
                  prior.scale = 1/sqrt(2),
                  prior.df = 1)
priorLine <- dcauchy(x = xx, location = -get_CohenH(k1, k3, n1, n3, NImargin), scale = 1/sqrt(2))
  

## plot posterior based on default prior

plot_df <- data.frame(xx, yy, priorLine)
names(plot_df) <- c("Effect Size", "Posterior", "Prior")

ggplot(plot_df, aes(x = xx)) + 
  geom_line(aes(y=yy, color = "Posterior"), size=1) + 
  geom_line(aes(y=priorLine, color = "Prior"), size = 1) + 
  labs(title = "Posterior Distribution with Default Prior", 
       subtitle = expression(paste(BF["+"]["-"],"(", d, "; ", 0.1059, ", ", 1/sqrt(2), ", ",1, ") = 28.87")),
       y = "Density", 
       x = "Effect size", 
       colour = "Distribution") + 
  scale_color_manual(values=c("#009E73","#0072B2")) +
  theme(legend.position = c(0.85, 0.75))

```

Reanalysis Comparison:

```{r}
#beta-lactam vs beta-lactam-macrolide
NI_BFprop (n1, n2, k1, k2, NImargin, 
           prior.location = -get_CohenH(k1, k2, n1, n2, NImargin), 
           prior.scale = 1/sqrt(2))     # Non-inferiority BF = 1307.764

#beta-lactam vs beta-lactam-fluoroquinolone
NI_BFprop (n1, n3, k1, k3, NImargin, 
           prior.location = -get_CohenH(k1, k3, n1, n3, NImargin), 
           prior.scale = 1/sqrt(2))     # Non-inferiority BF = 39.07749
```
The values that resulted do not match what was reported (1307.76 and 39.07, respectively) in the Ravenzwaaij paper. 


## Visualizations for Report


### Frequentist Analysis for beta-lactam vs beta-lactam-fluoroquinone

```{r}


n1 <- 656  # n for group 1 (beta-lactam)
n2 <- 739  # n for group 2 (beta-lactam-macrolide)
n3 <- 888 # n for group 1 (fluoroquinolone)
k1 <- 59  # mortality count in group 1 (beta-lactam)
k2 <- 82  # mortality count in group 2 (beta-lactam-macrolide)
k3 <- 78  # mortality count in group 1 (fluoroquinolone)

p1 <- k1/n1
p2 <- k2/n2
p3 <- k3/n3

NImargin <- 0.03   # Non-inferiority margin in proportion points

#perform approximate Z test to see if difference in treatment and control proportions is < - M

PooledProp <- (k1+k3) / (n1+n3)

#calculate z-stat for testing difference in 2 population proportions. 
# H_0: p3 -p1 > -0.03

Z <- ((p1-p3) - 0.03)/(sqrt(PooledProp*(1-PooledProp)*((1/n1) + (1/n3))))

pnorm(Z)

# since p-value is significant, this indicates strong evidence in favor of the H_A, which is support that drugs are NI

# calculate 95% CI

z_ref <- c(1.645, 1.96)

CI <- c("90%", "95%")

Group <- c("BM v B", "BM v B", "F v B", "F v B" )

LB <- c((p2-p1) - z_ref[1]*sqrt((p1*(1-p1)/n1) + (p2*(1-p2)/n2)), 
        (p2-p1) - z_ref[2]*sqrt((p1*(1-p1)/n1) + (p2*(1-p2)/n2)),
        (p3-p1) - z_ref[1]*sqrt((p1*(1-p1)/n1) + (p3*(1-p3)/n3)),        
        (p3-p1) - z_ref[2]*sqrt((p1*(1-p1)/n1) + (p3*(1-p3)/n3)))

UB <- c((p2-p1) + z_ref[1]*sqrt((p1*(1-p1)/n1) + (p2*(1-p2)/n2)), 
        (p2-p1) + z_ref[2]*sqrt((p1*(1-p1)/n1) + (p2*(1-p2)/n2)),
        (p3-p1) + z_ref[1]*sqrt((p1*(1-p1)/n1) + (p3*(1-p3)/n3)),
        (p3-p1) + z_ref[2]*sqrt((p1*(1-p1)/n1) + (p3*(1-p3)/n3)))

pt_est <- c((p2-p1), (p2-p1), (p3-p1), (p3-p1))



CI_data <- data.frame(Group, z_ref, CI, pt_est, LB, UB)

#Create the NI CI plot

ggplot(data=CI_data, aes(y=nrow(CI_data):1)) +
                geom_errorbar(aes(xmax = UB, xmin = LB, color=Group)) +
                geom_vline(xintercept = -0.03, 
                           color="#661100", 
                           lwd=0.5, 
                           linetype="dashed") +
                geom_point(aes(y=nrow(CI_data):1, x = pt_est), shape=1, size=3) +
                labs(title = "NI Confidence Interval Analysis",
                     y = "", x = "Effect Size (Treatment v Beta-Lactam)") +
                geom_text(aes(label = CI, colour = Group, 
                              x = UB + 0.0005, y = nrow(CI_data):1), hjust = -.1, size = 3) + 
                scale_color_manual(values=c("#009E73","#0072B2")) +
                theme(axis.text.y = element_blank(), axis.ticks = element_blank())
   

```

The negative point estimate favors the fluroquinone treatment. The 90% CI is completely above prespecified -c, so we can say that NI of fluoroquinone is demonstrated


### Prior vs Posterior Distribution Plots


Simulation for varying prior parameters

```{r}

n1 <- 656  # n for group 1 (beta-lactam)
n3 <- 888  # n for group 1 (fluoroquinolone)
k1 <- 59   # mortality count in group 1 (beta-lactam)
k3 <- 78   # mortality count in group 3 (fluorquinonlone)

prior.location <- seq(-1, 1, by = 0.01)
prior.scale <- seq(0.01, 1, by = 0.01)

NImargin <- 0.03

simulation_params <- list(
  prior.location = prior.location, 
  prior.scale = prior.scale
)

df <- cross_df(simulation_params)

df_BF <- df %>% 
    mutate(
    BF = unlist(pmap(list(k1 = k1, k2=k3, n1=n1, n2 = n3, 
                          NImargin = NImargin, 
                          prior.location = prior.location, 
                          prior.scale = prior.scale), NI_BFprop))
    ) 
  
summary(df_BF)

```


```{r}

ggplot(data = df_BF, aes(x = prior.location, y = prior.scale))+
  metR::geom_contour2(aes(z = log(BF), 
                          label = stat(level)), 
                      label_size = 3,
                      label_color = "#6699CC",
                      breaks = seq(-3, 10, 0.5)) + 
  labs(title = "Contour Plot of log(BF)",
       y = "Prior Scale", 
       x = "Prior Location")

```

Simulation for varying mortality counts


```{r}

#parameters for varying mortality count
k1 = seq(1, 0.25*n1, by = 10)
k3 = seq(1, 0.25*n2, by = 10)
 
NImargin <- 0.03
n1 <- 656
n3 <- 888
  
simulation_params2 <- list(
  k1 = k1,
  k3 = k3
)

#cannot take sqrt(negative numbers), so set the CohenH value to 0 if result is NaN

df2 <- cross_df(simulation_params2) %>% 
  mutate(CohenH = ifelse(is.na(get_CohenH(k1, k3, n1, n3, NImargin)), 0, get_CohenH(k1, k3, n1, n3, NImargin))
         )

df_BF2 <- df2 %>% 
  mutate(
    BF = unlist(pmap(list(k = k1, k2 = k3, n1=n1, n2 = n3, 
                          NImargin = NImargin, 
                          prior.location = -CohenH, 
                          prior.scale = 1/sqrt(2)), NI_BFprop)), 
    delta = (k3/n3)- (k1/n1), 
    Z = (((k3/n3) - (k1/n1)) + NImargin)/sqrt(((k1/n1)*(1-(k1/n1))/n1) + ((k3/n3)*(1-(k3/n3))/n3))
    ) 
  
summary(df_BF2)

```


Visual plots


```{r}


ggplot(data = df_BF2, aes(x = k1, y = k3))+
  metR::geom_contour2(aes(z = log(BF), 
                          label = stat(level)), 
                      label_size = 3,
                      label_color = "#332288",
                      breaks = seq(-20, 20, 3)) +
    metR::geom_contour2(aes(z = delta, 
                          label = stat(level)),
                        color = "#117733",
                      label_size = 3,
                      label_color = "#117733",
                      breaks = seq(-1, 1, 0.05))+ 
  labs(title = "Contour Plot of log(BF)",
       y = "Mortality Count in Fluoroquinolone Group", 
       subtitle = "Overlapped with Contour Plot of Effect Size in Green",
       x = "Mortality Count in Beta-Lactam Group")

ggplot(data = df_BF2, aes(x = k1, y = k3))+
  metR::geom_contour2(aes(z = log(BF), 
                          label = stat(level)), 
                      label_size = 3,
                      label_color = "#332288",
                      breaks = seq(-20, 20, 3)) +
    metR::geom_contour2(aes(z = Z, 
                          label = stat(level)),
                        color = "#882255",
                      label_size = 3,
                      label_color = "#882255",
                      breaks = seq(-6, 6, 2))+ 
  labs(title = "Contour Plot of log(BF)",
       y = "Mortality Count in Fluoroquinolone Group",
       subtitle = "Overlapped with Contour Plot of Z-Statistic in Purple", 
       x = "Mortality Count in Beta-Lactam Group")
  
```

Fill contour based on Frequentist and BF decision for NI

```{r}
ggplot(data = df_BF2, aes(x = k1, y = k3))+
  metR::geom_contour2(aes(z = log(BF), 
                          label = stat(level)), 
                      label_size = 3,
                      label_color = "#332288",
                      breaks = seq(-20, 26, 2)) +
  metR::geom_contour2(aes(z = Z, 
                          label = stat(level)),
                        color = "#882255",
                      label_size = 3,
                      label_color = "#882255",
                      breaks = seq(-12, 8, 2))+ 
  metR::geom_contour_fill(aes(z = Z), 
                          breaks = seq(2, 10), 
                          alpha = 0.2, 
                          linetype = "blank") +
  labs(title = "Contour Plot of log(BF)",
       y = "Mortality Count in Treatment Group",
       subtitle = "Overlapped with Contour Plot of Z-Statistic in Purple", 
       x = "Mortality Count in Control Group") +
  theme(legend.position="none")

ggplot(data = df_BF2, aes(x = k1, y = k3))+
  metR::geom_contour2(aes(z = log(BF), 
                          label = stat(level)), 
                      label_size = 3,
                      label_color = "#332288",
                      breaks = seq(-20, 26, 2)) +
  metR::geom_contour2(aes(z = Z, 
                          label = stat(level)),
                        color = "#882255",
                      label_size = 3,
                      label_color = "#882255",
                      breaks = seq(-12, 8, 2)) + 
  metR::geom_contour_fill(aes(z = log(BF)), 
                          breaks = seq(1, 26), 
                          alpha = 0.2, 
                          linetype = "blank") +
  labs(title = "Contour Plot of log(BF)",
       y = "Mortality Count in Treatment Group",
       subtitle = "Overlapped with Contour Plot of Z-Statistic in Purple", 
       x = "Mortality Count in Control Group") +
  theme(legend.position="none")


```
Zooming into the lower left corner

```{r}

ggplot(data = df_BF2, aes(x = k1, y = k3))+
  metR::geom_contour2(aes(z = log(BF), 
                          label = stat(level)), 
                      label_size = 3,
                      label_color = "#332288",
                      breaks = seq(-20, 20, 3)) +
  scale_fill_gradientn(colours = c("blue","blue", "blue", "blue", "blue")) +
  metR::geom_contour_fill(aes(z = Z),
                      breaks = seq(2,6),
                      alpha = 0.2, 
                      linetype = "blank") +
  labs(title = "Decision Boundaries on Contour Plot of log(BF)",
       y = "Mortality Count in Fluoroquinolone Group",
       x = "Mortality Count in Beta-Lactam Group") +
  theme(legend.position="none")

```

Zooming into the lower left corner, NO coloring

```{r}

ggplot(data = df_BF2, aes(x = k1, y = k3))+
  metR::geom_contour2(aes(z = log(BF), 
                          label = stat(level)), 
                      label_size = 3,
                      label_color = "#332288",
                      breaks = seq(-20, 20, 3)) +
  scale_fill_gradientn(colours = c("#42F0D9","#42F0D9", "#42F0D9", "#42F0D9"))+
  metR::geom_contour_fill(aes(z = log(BF)), 
                          breaks = seq(0, 20), 
                          alpha = 0.3, 
                          linetype = "blank") +
  labs(title = "Decision Boundaries on Contour Plot of log(BF)",
       y = "Mortality Count in Fluoroquinolone Group",
       x = "Mortality Count in Beta-Lactam Group") +
  theme(legend.position="none")

```

Change the breaks

```{r}

ggplot(data = df_BF2, aes(x = k1, y = k3))+
  metR::geom_contour2(aes(z = log(BF), 
                          label = stat(level)), 
                      label_size = 3,
                      label_color = "#332288",
                      breaks = seq(-20, 24, 3)) +
  scale_fill_gradientn(colours = c("#42F0D9","#42F0D9", "#42F0D9", "#42F0D9"))+
  metR::geom_contour_fill(aes(z = log(BF)), 
                          breaks = seq(0, 22), 
                          alpha = 0.3, 
                          linetype = "blank") +
  labs(title = "Contour Plot of log(BF) + Decision Boundaries",
       y = "Mortality Count in Treatment Group",
       subtitle = "Colored Decision Boundaries", 
       x = "Mortality Count in Control Group") +
  theme(legend.position="none")


```



Plot Instructional Visuals of Posterior vs Prior Distributions


```{r}
#create function to automate ggplots

post_plot <- function(prior.location, prior.scale){
  xlim <- range(c(-1, 1))
  xx <- seq(xlim[1], xlim[2], length.out = 400)
  yy <- posterior_t(delta = xx, t = Z, n1 = n3, n2 = n1,
                  independentSamples = TRUE,
                  prior.location = -get_CohenH(k1, k3, n1, n3, NImargin), 
                  prior.scale = 1/sqrt(2),
                  prior.df = 1)

  priorLine <- dcauchy(x = xx, location = prior.location, scale = prior.scale)
  
  plot_df <- data.frame(xx, yy, priorLine)
  names(plot_df) <- c("Effect Size", "Posterior", "Prior")

  gplot <- ggplot(plot_df, aes(x = xx)) + 
    geom_line(aes(y=yy, color = "Posterior"), size=1) + 
    geom_line(aes(y=priorLine, color = "Prior"), size = 1) +
    ylim(0, 8)+
    labs(y = "Density",
        x = "Effect size",
        subtitle = paste("Prior loc:", prior.location,", Prior scale:", prior.scale)) + 
    scale_color_manual(values=c("#009E73","#0072B2")) +
    theme(legend.position="none", 
          text = element_text(size = 8))
  
  return(gplot)
}


n1 <- 656  # n for group 1 (beta-lactam)
n3 <- 888  # n for group 1 (fluoroquinolone)
k1 <- 59   # mortality count in group 1 (beta-lactam)
k3 <- 78   # mortality count in group 3 (fluorquinonlone)

SE <- sqrt (((k1/n1)*(1-(k1/n1))/n1) + ((k3/n3)*(1-(k3/n3))/n3))
  
Z <- (((k3/n3) - (k1/n1)) + NImargin) / SE

prior.location <- c(-0.75, 0, 0.5)
prior.scale <- c(0.05, 0.15, 0.5)

simulation_params3 <- list(
  prior.location = prior.location, 
  prior.scale = prior.scale
)

df3 <- cross_df(simulation_params3)

p <- list()

for(i in 1:nrow(df3)){
  p[[i]] <- post_plot(df3$prior.location[i], df3$prior.scale[i])
}


#create a plot so that the legend can be extracted out
plot_legend <- post_plot(df3$prior.location[1], df3$prior.scale[1])+
              theme(legend.position = "right", )

legend <- get_only_legend(plot_legend)

do.call(grid.arrange,list(grobs = p, 
                          top = "Posterior vs Prior Comparison", 
                          right = legend))



```
